Dependently Typed Languages in Statix

### 2023-03-17 pres-evcs/presentation.tex
add note
Update presentation

\begin{frame}[fragile]{Research Question}
\large{How suitable is Statix for defining a dependently-typed language?}
% Statix is turing complete, so it's going to be possible
% But is it going to be easier or harder than a general purpose language like haskell?
\end{frame}

@@

\begin{frame}{Primary Contribution}
	todododododo
\end{frame}

\begin{frame}[fragile]{Calculus of Constructions}
	A lambda calculus with dependent types.

@@

\begin{frame}[fragile]{Type Checking}
	\begin{block}{Type checking relation}
		\texttt{typeOfExpr : scope * Expr -> Expr}
	\end{block}
	% Returns an Expr, not a Type
	% Scope -> Two different types of nodes
	\begin{block}{How do we use scopes?}
		A scope is used as a combination of an environment and a context.\\
		One relation \texttt{name $\rightarrow$ NameEntry}, NameEntry is either:
		\begin{itemize}
			\item NType: Stores a type -> Corresponds with context
			\item NSubst: Stores a substitution -> Corresponds with environment
		\end{itemize}
	\end{block}
\end{frame}

@@

\begin{frame}[fragile]{Type Checking: From inference rules to Statix code}
	todododododod From inference rules to Statix code

	\begin{figure}
		\centering
		\includegraphics[width=0.5\linewidth]{img/screenshot001}
		\label{fig:screenshot001}
	\end{figure}
	\begin{figure}
		\centering
		\includegraphics[width=0.5\linewidth]{img/screenshot002}
		\label{fig:screenshot002}
	\end{figure}

@@

\begin{frame}[fragile]{Conclusions}
	Spoofax is a great tool for developing dependently typed languages!\footnote{But there is still room for improvement}

	- using scopes as env+context 
	todo beperkingen noemen 
	- statix enforces syntactic uniqueness of metavariable solutions but dependently typed languages use a weaker notion of equality 
	- full thesis coming soonTM
\end{frame}

>>>>

\begin{frame}[fragile]{Research Question}
\large{How suitable is Statix for defining a dependently-typed language?}
\begin{itemize}
	\item Defining system F was a challenge\footnote{Hendrik van Antwerpen et al. Scopes as types.}, will it even be possible?
	\item Will it be easier than doing it in Haskell?
\end{itemize}

% Mention system f (polymorphic lambda calculus) being hard (scopes as types)
% Statix is turing complete, so it's going to be possible
% But is it going to be easier or harder than a general purpose language like haskell?
\end{frame}

@@

\begin{frame}[fragile]{Primary Contribution: Calculus of Constructions in Statix}
	A lambda calculus with dependent types.

@@

\begin{frame}[fragile]{Type Checking}
	\begin{block}{Type checking relation}
		\texttt{typeOfExpr : scope * Expr -> Expr}
	\end{block}
	% Returns an Expr, not a Type
	% Scope -> Two different types of nodes
	\begin{block}{How do we use scopes?}
		A scope is used as a replacement for an environment and a context..\\
		One relation \texttt{name $\rightarrow$ NameEntry}, NameEntry is either:
		\begin{itemize}
			\item NType: Stores a type (Corresponds to a context)
			\item NSubst: Stores a substitution (Corresponds to an environment)
		\end{itemize}
	\end{block}
\end{frame}

@@

\begin{frame}[fragile]{Type Checking: From inference rules to Statix code}
\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{img/screenshot004}
\end{figure}

\begin{figure}
	\centering
	\begin{exampleblock}{Equivalent Statix code} 
		\texttt{typeOfExpr (s, Let(x, e, b)) = bt :-\\
			\,\, typeOfExpr (s, e) == et, \\
			\,\, typeOfExpr (sPutSubst (s, x, (s, e)), b) == bt}
	\end{exampleblock}
\end{figure}

@@

\begin{frame}[fragile]{Conclusions}
	Spoofax is a great tool for developing dependently typed languages!
	\begin{itemize}
		\item We can use scopes to represent environments and contexts
		\item Statix can still use improvements
	\end{itemize}

	%- using scopes as env+context  
	%- statix enforces syntactic uniqueness of metavariable solutions but dependently typed languages use a weaker notion of equality 
	%- full thesis coming soonTM
\end{frame}
